rawPaper

### 2020-10-23 analysis/paper/rawPaper.Rmd
names & length set --- pending alignment

```{r, echo=FALSE}
chunk for generating the map.
```

@@

# checking/setting working directory as raw_data for storing the final fasta file, which are the sequences before aligning them in the next chunk 
getwd()
# setwd is temporal, after fishining this chunk, the wd will be again /analysis/paper
setwd("/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data")

@@

# storing ALL sequences in a single fasta file
write(Sicydium_COI_seqs1, "Sicydium_COI_seqs.fasta", sep = "\n")
write(Sicydium_COI_seqs2, "Sicydium_COI_seqs.fasta", sep = "\n", append = TRUE)   # by setting 'append = TRUE', I can keep adding seqs
write(Awaous_bananaCOI_seqs, "Sicydium_COI_seqs.fasta", sep = "\n", append = TRUE)
write(Awaous_ocellarisCOI_seqs, "Sicydium_COI_seqs.fasta", sep = "\n", append = TRUE)

@@

# reading the file and renaming the sequences
Sicydium_COI_seqinrFormat <- read.fasta(file = "Sicydium_COI_seqs.fasta", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)

@@

# writing a new fasta file using the vector above 
write.fasta(sequences = Sicydium_COI_seqinrFormat, names = Sicydium_COI_GenBank_Ids, nbchar = 10, file.out = "Sicydium_COI_seqinrFormat.fasta")

@@

# removing all sequences without a complete scientific name, that is, those with 'sp.'
fastafile <- read.fasta(file = "Sicydium_COI_seqinrFormat.fasta", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)

getNames <- strsplit(names(fastafile), '_')         # creating a vector splitting species names
removeSeqs <- c("sp.")                              # finding names to remove: indexes which are in the species to be removed
keepSeqs <- which(! getNames %like% removeSeqs)     # selection those without specific name in them

# saving a new fasta file with only those seqs with full scientific names
write.fasta(sequences=fastafile[keepSeqs], names=names(fastafile)[keepSeqs], file.out="Sicydium_COI_edits.fasta")

>>>>

```{r, echo=FALSE}
#chunk for generating the MAP.

install.packages("maps")
install.packages("mapdata")

library(maps)
library(mapdata)
map("worldHires","Cuba", xlim=c(-88,-73), ylim=c(19,24), col="gray95", fill=TRUE)

@@

# checking/setting working directory as raw_data for storing the final fasta file, which are the sequences before aligning them in the next chunk 
#getwd()
# setwd is temporal, after fishining this chunk, the wd will be again /analysis/paper
#setwd("/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data")
#getwd()

@@

# storing ALL sequences in a single fasta file
write(Sicydium_COI_seqs1, file="../data/raw_data/Sicydium_COI_seqs.fasta", sep = "\n")
write(Sicydium_COI_seqs2, file="../data/raw_data/Sicydium_COI_seqs.fasta", sep = "\n", append = TRUE)   # 'append = TRUE', I can keep adding seqs
write(Awaous_bananaCOI_seqs, file="../data/raw_data/Sicydium_COI_seqs.fasta", sep = "\n", append = TRUE)
write(Awaous_ocellarisCOI_seqs, file="../data/raw_data/Sicydium_COI_seqs.fasta", sep = "\n", append = TRUE)

@@

# reading the file and renaming the sequences
Sicydium_COI_seqinrFormat <- read.fasta(file ="../data/raw_data/Sicydium_COI_seqs.fasta", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)

@@

# writing a new fasta file (2nd) using the vector above 
write.fasta(sequences = Sicydium_COI_seqinrFormat, names = Sicydium_COI_GenBank_Ids, nbchar = 10, file.out = "../data/raw_data/Sicydium_COI_seqinrFormat")

@@

### Removing sequences of incomplete/unspecific NAMES; that is, those with 'sp.'
nameFasta <- read.fasta(file = "../data/raw_data/Sicydium_COI_seqinrFormat", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)

fastaNames <- strsplit(names(nameFasta), '_')                  # creating a vector splitting species names
completeSeqs <- which(! grepl("sp.", fastaNames))              # selecting those with specific name in them       # 562

# writing a new fasta (3rd) file with only those seqs with full scientific names
write.fasta(sequences = nameFasta[completeSeqs], names = names(nameFasta)[completeSeqs], file.out = "../data/raw_data/Sicydium_COI_names.fasta")

### Removing sequences of LENGTH shorter than 600 pb, given that the gene COI has between 627 and 658 bp.
lengthFasta <- read.fasta(file = "../data/raw_data/Sicydium_COI_names.fasta", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)
fastaLength <- lengthFasta[which(getLength(lengthFasta) >= 600)]       # 399
write.fasta(sequences = fastaLength, names = names(fastaLength), file.out = "../data/raw_data/Sicydium_COI_lengths.fasta")

### Aligning the sequences

BiocManager::install("msa")
library(msa)

alignmentFasta <- read.fasta(file = "../data/raw_data/Sicydium_COI_lengths.fasta", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)

length(alignmentFasta)

myClustalOmegaAlignment <- msa("alignmentFasta", type = "DNA", "ClustalOmega")

print(myClustalOmegaAlignment, showConsensus=FALSE, halfNrow=3)
print(myClustalOmegaAlignment, showNames=FALSE, show="complete")

### Removing most duplicated sequences

### Removing the segment before the start codon TAC
